A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2020-08-27 Papers/Amlap_2020_talk/amlap_2020_talk.tex
work on talk

\begin{frame}
\maketitle
%TODO maybe put a picture of Roger's face??, at least acknowledge him 
\end{frame}

\begin{frame}{Incremental processing methods}
\pause
Want a measure of processing difficulty \greencheck\pause

@@

			\includegraphics[width=.9\textwidth]{../eye-tracker.jpeg}
			\pause
		\end{center}
	\end{column}\pause
	\begin{column}{.5\textwidth}
		\begin{center}
\textbf{\large Self-paced reading}

@@

\pause 
\centering

@@

	Want to use Maze instead of web SPR\pause

@@

	Needs some tweaks:\pause
	\begin{itemize}
		\item Run on web \pause
		\item Easily generate distractors \pause
		\item Work for multi-sentence items \pause
	\end{itemize}

@@

\begin{frame}{Run on web}
	Wrote an Ibex module \pause

@@

	Want to use Maze instead of web SPR

@@

LSTM RNNs show some syntactic knowledge %TODO cite people 
\pause

@@

Problem: Errors terminate sentences. \pause
\begin{itemize}
	\item Treat whole story as a unit: ~ No one makes it to the end. \pause
	\item Treat each sentence as a unit: Some participants miss key context. \pause
\end{itemize}

@@

	\end{column} \pause 
	\begin{column}{0.6\textwidth} 
		\begin{center}
			\begin{itemize}
				\item Can be toggled in Ibex \pause
				\item Long materials feasible \pause
				\item Have all the data \pause
				\item Compensates for bad distractors
			\end{itemize}
		\end{center}
	\end{column}
\end{columns}
\medskip
\pause
\end{frame}
%

@@

\begin{frame}{Natural Stories}
Natural stories corpus (Futrell et al. 2017) \pause
\begin{itemize}
	\item 10 stories, each about 1000 words \pause
	\item Some unusual constructions, but read fluently \pause
	\item 6 comprehension questions per story \pause
\end{itemize}

@@

@@

%HERE%
\begin{frame}{Surprisal Effects}
%TODO copy from updated abstract
Use 3 LMs to estimate surprisal: smoothed 5-gram, Gulordava RNN, Transformer-XL (Gulordava et al. 2018, Dai et al. 2019)
\pause

%TODO make sure that updated correct gams are used!!!

Fit GAMs on pre-error data.
\begin{itemize}
	\item Limit to single-token words.
	\item Fit to both current and past word surprisal.
\end{itemize}

@@

\begin{frame}
	Implications for surprisal theory -- size of the effects is different
	include effect sizes
	Is it about participant quality?
\end{frame}

\begin{frame}{Surprisal Effects}
\includegraphics[width=.9\textwidth]{../gam.pdf}	
\end{frame}
\begin{frame}{What about post-mistake data?}
Exclude data from mistakes or the two words after a mistake.

\includegraphics[width=.9\textwidth]{../gam2.pdf}	
\end{frame}

\begin{frame}{Summary}
\begin{itemize}\pause
	\item People will read in the Maze task for 15-20 minutes \pause
	\item It's possible to comprehend during Maze \pause
	\item Distractors are generally good enough \pause
	\item Find expected RT patterns \pause
	\item Very little spillover
\end{itemize}
\end{frame}

\begin{frame}{Use A-maze!}
Easy to use! \pause
\begin{itemize}
	\item Runs on command line \pause
	\item Match distractors across minimal pair sentences \pause
	\item Customize surprisal thresholds, vocabulary lists \pause
	\item Can output pre-formatted for Ibex
\end{itemize}
\medskip
\end{frame}

\begin{frame}{Use A-maze!} \pause
Contribute to A-maze:
\begin{itemize}
	\item Written in Python 3 \pause
	\item Interface with other language models \pause
	\item Add more frequency sources \pause
	\item Extend to non-English languages
\end{itemize}

@@

\begin{frame}{}

\textcolor{ForestGreen}{\large Documentation: vboyce.github.io/Maze}

with links to the following:
\begin{itemize}

\item A-maze code: github.com/vboyce/Maze

\item Web-maze code: github.com/vboyce/Ibex-with-Maze

\item Sample task: syntaxgym.org:666

\item Paper: psyarxiv.com/b7nqd/
\end{itemize}
\end{frame}

@@

\begin{frame}{Matching distractors}
If unspecified: Match by position
\begin{itemize}
	\item The son of the lady who politely introduced \sethlcolor{green} \hl{herself}\sethlcolor{pink} / \hl{himself} was popular at the party.
\end{itemize}
Can specify labels for each word to pair (within item)
\begin{itemize}
	\item The cat who the dog scared hid in a box.\\pre-1 pre-2 who art noun verb main-verb post-1 post-2 post-3
\item The dog who scared the cat sniffed around the couch.\\ pre-1 pre-2 who verb art noun main-verb post-1 post-2 post-3
\end{itemize}
\end{frame}
%%back pocket on more on model results??
%%do we even need back pockets??

\begin{frame}{Post-mistake only}
Exclude data from mistakes or the two words after a mistake.

\includegraphics[]{../gam3.pdf}	
\end{frame}

\begin{frame}{Regression coefficients}
\begin{tiny}
\begin{tabular}{l|rlr|rlr|rlr}
	\hline
	&\multicolumn{3}{c|}{5-gram}&\multicolumn{3}{c|}{GRNN}&\multicolumn{3}{c}{TXL}\\
	& Est & CI & $p$ & Est & CI & $p$ &Est & CI & $p$ \\ 
	\hline
	Intercept & 865.3 & [829.9, 902.9] & 0.00 & 871.1 & [837.9, 905.3] & 0.00 & 870.8 & [832.5, 907.8] & 0.00 \\ 
	Surprisal & 11.7 & [9.3, 14.1] & 0.00 & 23.7 & [21, 26.5] & 0.00 & 18.5 & [16.1, 21.1] & 0.00 \\ 
	Frequency & -2.9 & [-6.3, 0.5] & 0.10 & 2.9 & [-0.2, 6] & 0.06 & 0.4 & [-2.7, 3.5] & 0.79 \\ 
	Length & 20.5 & [15.4, 25.6] & 0.00 & 18.5 & [13.3, 23.7] & 0.00 & 21.4 & [16.2, 26.6] & 0.00 \\ 
	Surprisal:Length & -2.0 & [-3, -1] & 0.00 & -1.8 & [-2.7, -0.9] & 0.00 & -1.4 & [-2.2, -0.6] & 0.00 \\
	Freq:Length & -1.0 & [-2.5, 0.4] & 0.16 & -0.1 & [-1.2, 1] & 0.82 & 0.2 & [-0.9, 1.2] & 0.76 \\ 
	\hline
	Past Surprisal & 1.6 & [-0.5, 3.6] & 0.14 & 2.7 & [0.8, 4.5] & 0.00 & 0.8 & [-0.9, 2.5] & 0.40 \\ 
	Past Freq & 2.6 & [-0.1, 5.4] & 0.06 & 1.9 & [-0.2, 4.2] & 0.08 & 1.2 & [-1.1, 3.6] & 0.30 \\ 
	Past Length & -4.8 & [-9, -0.1] & 0.04 & -6.6 & [-10.9, -2.1] & 0.00 & -5.2 & [-9.3, -0.7] & 0.03 \\ 
	Past Surp:Length & -0.2 & [-1.2, 0.8] & 0.72 & -0.9 & [-1.7, -0.2] & 0.01 & -0.6 & [-1.3, 0.2] & 0.13 \\ 
	Past Freq:Length & -1.0 & [-2.3, 0.3] & 0.15 & -1.8 & [-2.9, -0.8] & 0.00 & -1.5 & [-2.6, -0.5] & 0.01 \\

	\hline
\end{tabular}
\end{tiny}
\end{frame}

>>>>

\begin{frame}{ }
\vskip 6em

\begin{center}
{\LARGE A-maze of Natural Stories:}

\smallskip

{\Large Texts are comprehensible using\\ \smallskip the Maze task}

{\large Veronica Boyce, Roger Levy}

{\large AMLaP 2020}

\begin{tikzpicture}[remember picture,overlay]
\node[xshift=-1.8cm,yshift=-1cm] at (current page.east) {\includegraphics[clip, trim=.1cm 0cm .2cm 0cm,width=.2\textwidth]{roger.jpg}};
\end{tikzpicture}

\begin{tikzpicture}[remember picture,overlay]
\node[xshift=2.5cm,yshift=-1.5cm] at (current page.north west) {\includegraphics[width=.3\textwidth]{cpl_header.png}};
\end{tikzpicture}

%TODO maybe put a picture of Roger's face??, at least acknowledge him 
\end{center}
\end{frame}

\begin{frame}{Incremental processing methods}
\pause
Want a measure of processing difficulty\pause

@@

			\includegraphics[width=.9\textwidth]{../eye-tracker.jpeg}
			\pause
		\end{center}
	\end{column}
	\begin{column}{.5\textwidth}
		\begin{center}
\textbf{\large Self-paced reading}

@@

\centering

@@

	Can we use Maze instead of web SPR?\pause

@@

	Needs some tweaks:\pause
	\begin{itemize}
		\item Run on web \pause
		\item Easily generate distractors \pause
		\item Work for multi-sentence items
	\end{itemize}

@@

\begin{frame}{Run on web}

	\pause
	Wrote an Ibex module \pause

@@

	Can we use Maze instead of web SPR?

@@

Run items through LM, choose high surprisal words as distractors
\pause

@@

Problem: Errors terminate sentences. \pause
\begin{itemize}
	\item Treat whole story as a unit: \pause ~ No one makes it to the end. \pause
	\item Treat each sentence as a unit: \pause Some participants miss key context. \pause
\end{itemize}

@@

	\end{column} \pause 
	\begin{column}{0.6\textwidth} 
		\begin{center}
			\begin{itemize}
				\item Can be toggled in Ibex Maze \pause
				\item Long materials feasible \pause
				\item Have all the data \pause
				\item Compensates for bad distractors
			\end{itemize}
		\end{center}
	\end{column}
\end{columns}
\end{frame}
%

@@

\begin{frame}{Natural Stories}
Natural stories corpus (Futrell et al. 2017) \pause
\begin{itemize}
	\item 10 stories, each about 1000 words \pause
%	\item Some unusual constructions, but read fluently \pause
	\item 6 comprehension questions per story
\end{itemize}

@@

\pause

@@

%HERE%
\begin{frame}{Surprisal Effects}
\large{Is RT linear in terms of surprisal?}\pause
\medskip

Estimate surprisal from 3 models:
\begin{itemize}
	\item smoothed 5-gram
	\item LSTM-RNN (Gulordava et al. 2018)
	\item Transformer-XL (Dai et al. 2019)
\end{itemize}

\pause

Fit GAMs
\begin{itemize}
	\item Limit to single-token words
	\item Fit to both current and past word surprisal
	\item Include frequency, length as predictors
\end{itemize}

@@

\begin{frame}{Surprisal Effects}
	\includegraphics[width=.9\textwidth]{../gam.pdf}	
\end{frame}

\begin{frame}{Surprisal Effects}
	Linear Models \\
	\pause

	\begin{small}
		\begin{tabular}{l|r|r|r}
			\hline
			&5-gram & GRNN & TXL\\
			\hline
			Intercept & \textbf{865.3} & \textbf{871.1} & \textbf{870.8} \\ 
			Surprisal & \textbf{11.7}  & \textbf{23.7} & \textbf{18.5}  \\ 
			Frequency & -2.9  & 2.9  & 0.4  \\ 
			Length & \textbf{20.5} & \textbf{18.5}  & \textbf{21.4} \\ 
			Surprisal:Length & \textbf{-2.0}  & \textbf{-1.8}  & \textbf{-1.4} \\
			Freq:Length & -1.0  & -0.1 & 0.2  \\ 
			\hline
			Past Surprisal & 1.6  & \textbf{2.7} & 0.8 \\ 
			Past Freq & 2.6  & 1.9  & 1.2  \\ 
			Past Length & \textbf{-4.8}  & \textbf{-6.6 }&\textbf{ -5.2} \\ 
			Past Surp:Length & -0.2 & \textbf{-0.9} & -0.6  \\ 
			Past Freq:Length & -1.0  & \textbf{-1.8 }&   \textbf{-1.5}  \\ 
			\hline
		\end{tabular}

	Surprisal in bits, Length in characters, \\
	Frequency in $log_2$ occurrences/billion words
	\end{small}

\end{frame}
\begin{frame}{Surprisal Effects}
	Takeaways: \pause
	\begin{itemize}
		\item Minimal frequency effects (consistent with Shain 2019) \pause
		\item Long RTs \pause
		\item Large effects of Length, Surprisal \pause
		\item Very little spillover \pause
	\end{itemize}

	Model comparison: GRNN is best, but TXL complementary
\end{frame}

\begin{frame}{Why such large effects?}
	\pause
	Framework: Look at words long enough to ID with some threshold of certainty \pause

Could hypothesize mechanisms for difference: \pause
\begin{itemize}
	\item Task leads to higher threshold \pause 
	\item Task demands reduce available resources for processing \pause
	\item Presence of second word 
\end{itemize}

@@

\begin{frame}{Conclusion}

	\pause

Consider A-maze! 
\begin{itemize}\pause
	\item Documentation:  \textcolor{ForestGreen}{vboyce.github.io/Maze} \pause
	\item Versatile \pause
	\item Low spillover \pause 
\end{itemize}

Natural Stories A-maze: \pause 
\begin{itemize}
	\item Participants comprehend what they read \pause
	\item Find linear, large surprisal effects
\end{itemize}

@@

\begin{frame}{Surprisal Effects}
	GAM if we only exclude mistakes (all participants, post-mistake data)
	\includegraphics[width=.9\textwidth]{../gam-any.pdf}	
\end{frame}
